Topological Equivariant Artist Model for Visualization Library Architecture

### 2022-05-17 paper/main.tex
q

The fiber category is equipped with a bifunctor $\otimes: \mathcal{\dfiber} \times {\dfiber}$. The bifunctor provides a method for combining fibers, thereby allowing for fields that contain multityped values. For example, RGB color can either be represented as three fiber fields $\dfiber_{red}, \dfiber_{green}, \dfiber_{blue}$ or a composite fiber field $\dfiber_{red} \otimes \dfiber_{green} \otimes \dfiber_{blue} = \dfiber_{RGB}$.

@@

A sheaf is a presheaf that satisfies the locality and gluing axioms \cite{bakerMathsSheaf}. 
\begin{definition}
\label{def:atct:sheaf:locality}
The locality axiom is that given a union of opensets $\mathscr{\opensetc} = \bigcup\limits_{i\in I} \opensetc_i$ and $\dsectionc^{a}, \dsectionc^{b} \in \sheafc(\mathscr{\opensetc})$,  if $\dsectionc^{a}\restriction_{\opensetc_i} = \dsectionc^{b}\restriction_{\opensetc_i}$ for each $\opensetc_i \in \opensetc$ then $\dsectionc^{a} = \dsectionc^{b}$
\end{definition}
This means that two sections in a sheaf are equal when they evaluate to the same values over all open sets in a collection. 
\begin{definition}
\label{def:atct:sheaf:gluing}
The gluing axiom is that given $\dsectionc^{i} \in \sheafc(\opensetc_i)$ such that ${\dsectionc^{i}}\restriction_{\opensetc_i\cap \opensetc_j} = {\dsectionc^{j}}\restriction_{\opensetc_i \cap \opensetc_j}$ for $\opensetc_i, \opensetc_j \in \mathscr{\opensetc}$, there exists $\dsectionc \in \sheafc(\mathscr{\opensetc})$ such that $\dsectionc\restriction_{\opensetc_i} = {\dsectionc^{i}}$. 
\end{definition}
This means we can construct a section $\dsection$ on the collection of open sets that is the union of sections defined on specific open sets. This expresses an equivalency between a concatenated representation of data and the distributed representation of the same data. In this model, we represent data and graphics as a sheaf; doing so allows us to express that data sources are expected to have the bookkeeping that a sheaf requires, meaning inclusions, restrictions, locality, and gluing must be preserved.

\subsubsection{Structure on Data}
\label{sec:atct:sheaves:measurments}
We define structure on data as the allowable \textcolor{action}{transformations on the bundle}, which we formalize as actions $\dfuncc$ on sections of that sheaf. Our construction of $\dfunc$ is a generalization of the classification of measurement scales by their mathematical structure \cite{stevensTheoryScalesMeasurement1946,leaFormalizationMeasurementScale, thomasMathematizationNotMeasurement2014}, which often serves as the basis of the equivariance described in \autoref{sec:related-work:equivariance}.

@@

\begin{equation}
  \label{eq:artist:natural_transform:inclusions}
  \begin{tikzcd}
    \opensetc_1 & \cgamma{\opensetc_{1}}{\dtotalc\restruction_{\opensetc_1}} \arrow[dd, "\iota^*"', color=set] \arrow[rr, "\vartistc_{\opensetc_1}", color=artist] &  & \cgamma{\opensetc_{1}}{\vindexpushc\gtotalc\restruction_{\opensetc_1}} \arrow[dd, "\iota^*", color=set] \\
      &  &  &  \\
    \opensetc_{2} \arrow[uu, "\iota", hook, color=base] & \cgamma{\opensetc_{2}}{\dtotalc\restruction_{\opensetc_2}} \arrow[rr, "\vartistc_{\opensetc_2}", color=artist]                        &  & \cgamma{\opensetc_{2}}{\vindexpushc\gtotalc\restruction_{\opensetc_2}}                      
    \end{tikzcd}
\end{equation}
 The diagram in \autoref{eq:artist:natural_transform} shows that restricting a set of outputs of an artist to a set of graphic sections over a subspace is equivalent to restricting the inputs to data sections over the same subspace. Since the input and output of the artist is always sections over the same subspace, this is why the by construction the artist preserves the bookkeeping provided through the use of sheaves.

@@

\begin{equation}
  \label{eq:artist:inout:diagram}
  \begin{tikzcd}[row sep=huge]
    \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} 
    \arrow[rr, "\vartistc", color=artist] 
    \arrow[d, "\equivc"', color=monoid] &  & 
    \imartist{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
    \arrow[d, "render"] 
    \arrow[lld, "\extractmc"', color=monoid, dashed] \\
    {\textcolor{set}{Hom}(\opensetc, \measurec)}  &  & visualization 
    \arrow[ll, "measure",]
    \end{tikzcd}
\end{equation}
so instead we introduce a measurement extraction function $\extractm = measure\circ render$ that decomposes the rendered output into measurable components and returns a map from the data index into the measurement space
\begin{equation}
  \label{eq:artist:actual}
  \extractmc:\gsectionc \mapsto (\opensetc \xrightarrow{\extractmc_{\gsectionc}} \measurec)
\end{equation}
such that $\extractmc_{\gsectionc}(\dbasepoint)$ returns a quantification of the measurable components $\measurec_{\dbasepointc}$ of the visual element generated for the data at that point $\gsectionc|_{\vindexprec(\dbasepointc)}$.

We also introduce a function \equivc\ that maps data to the measurement space directly 
\begin{equation}
\equivc: \dsectionc \mapsto (\opensetc \xrightarrow{\equivc_{\dsectionc}} \measurec)
\end{equation}
such that $\equivc_{\dsection}(\dbasepointc)$ is the expected set of measurements $\measure_{\dbasepoint}$. The pair of \textcolor{monoid}{verification functions} (\equivc, \extractmc) can be used to test that the expected encoding $\equivb_{\dsection}$ of the data matches the actual encoding $\extractm_{\gsection}$ 
\begin{equation}
  \label{eq:artist:verification}
    \equivc(\dsectionc)(\dbasepointc) = \extractmc(\vartistc(\dsectionc))(\dbasepointc) = \extractmc(\gsectionc\circ\vindexprec)(\dbasepointc)=\measurec_{\dbasepointc}
\end{equation}

An artist is equivariant when changes to expected and actual are equivariant. As introduced in \autoref{eq:atct:morphism:base}, the base space transformation \dfunch\ is invariant because $\dsection\restriction_{\openset} = \dsection\restriction_{\dfunch(\openset^{\prime})}$. This means that, for all points in the data $\dbasepoint \in \dbase$, the measurement should not change if only the base space is transformed 
\begin{equation}
  \label{eq:atct:equivarance:verify:base}
  \equivc(\vartistc(\dsectionc))(\dfunch(\dbasepointc^{\prime})) = \extractmc(\vartistc(\dsectionc))(\dbasepointc)
\end{equation}
On the other hand, a change in sections \autoref{eq:atct:morphism:fiber} induces an equivalent change in measurements
\begin{equation}
  \label{eq:atct:equivarance:verify:fiber}
  \equivc(\dfunctc(\dsectionc))(\dbasepoint) = \dfunctc_{\measure}(\extractmc(\vartistc(\dsectionc))(\dbasepoint))
\end{equation}
The change in measurements $\dfunctc_{\measure}$ is defined by the developer as the symmetry between data and graphic that the artist is expected to preserve.

@@

\begin{equation}
  \label{eq:artist:plus:verify}
  \begin{split}
  \vartistc_{a+b}(\dsectionc^{a+b}(\dbasepointc_{a})) \\ 
  = \vartistc_{a}(\dsectionc^{a}(\dbasepointc_{a})) = \vartistc_{b}(\dsectionc^{b}(\dbasepointc_{b}))
  \end{split}
\end{equation}
 for all $\dbasepointc_{a}, \dbasepointc_{b} \in \dbasec_{a}\bigsqcup\limits_{\dbasec_{c}} \dbasec_{b}$

@@

\begin{equation}
  \begin{split}
    \extractmc(\vartistc_{a\times b}(\dsectionc^{a\times b}(\dbasepointc)))\restriction_{dfiberc^{c}}\\ = 
    \extractmc(\vartistc_{a}(\dsectionc^{a}(\dbasepointc_{a})))\restriction_{\dfiber^{c}} = \extractmc(\vartistc_{b}(\dsectionc^{b}(\dbasepointc_{b})))\restriction_{\dfiber^{c}}
  \end{split}
\end{equation}

@@

The circle-line glyph in \autoref{fig:artist:operator} can be generated via a line plot and highlighted points in that plot, which is the addition case of two datasets with the same fields and different continuities. The same circle-line glyph can also be generated via the multiplication case as different fields of the same bundle. As mentioned in \autoref{sec:artist:operator:multiplication}, multiplication is implemented as addition of two copies of a base space. Although we have been discussing the trivial cases of adding observations or adding fields, this merging of artists in datasets can be generalized:
\begin{equation}
  \vartist(\cgamma{\bigsqcup_{i} \dbase^{i}}{\bigoplus_{i}\dtotalc^{i}}) = \sum_{i}
  \vartist_{i}(\cgamma{\dbase^{i}}{\dtotal^{i}}) 
\end{equation}

@@

\begin{figure}[h!]
  \label{fig:construction:xi}
  \includegraphics[width=1\columwidth]{deform_retract.png}
  \caption{The graphic base space $\gbase$ is collapsable to the line $\dbase$ such that every band $(\dbasepoint_i, [0,1])$ on $\gbase$ maps to corresponding point $\dbasepoint_i \in \dbase$ and the band $[0,1]$ determines which pixels are filled in to generate the shape of the line.\note{maybe put in a phantom line plot in s}}
\end{figure}
For example, as shown in \autoref{fig:construction:xi}, a line is 1D but is a 2D glyph on a screen; therefore the graphic space $\gbase$ is constructed by multiplying the base space $\dbase$ with an interval $[0,1]$. Because $\gbase$ is collapsible into $\dbase$, every band $(\dbasepoint_i, [0,1])$ corresponds to a point in the base space $\dbasepoint_i \in \dbase$. The first coordinate $\alpha=\dbasepoint_i$ provides a lookup to retrieve the associated visual variables. The second coordinate, which is a point in the interval $\beta=[0,1]$. Together they are a point $\gbasepoint=(\alpha,\beta) \in gbase$ in the graphic base space. This point $\gbasepoint$ is the input into the graphic section $\gsection(\gbasepoint)$ that is used to determine which pixels are colored, which in turn determines the thickness, texture, and color of the line.

\subsection{Transforming Data into Measurable Components}
The transformation $\vchannel$ is a map from data sections to graphic sections
\begin{equation}
\nu: \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \rightarrow \cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}}
\end{equation}
such that the sections project to the same point on the base space $\pi(\dtotal) = \pi(\vchannel(\dtotal))$. As with artists, $\vchannel$ are maps of sections such that the composition operators defined in \autoref{sec:artist:composition} can also act on transformers $\vchannel$.

A consequence of $\pi(\dtotal) = \pi(\vchannel(\dtotal))$ is that $\vchannelc$ can be constructed as a pointwise transformation such that  
\begin{equation}
  \label{eq:constrution:nu}
  \vchannelc: \dfiberc_{\dbasepointc} \rightarrow \vfiberc_{\dbasepoint}
\end{equation}
which means that means that a point in a single data fiber $\delement \in \dfiber_{\dbasepointc}$ can be mapped into a corresponding point in a visual fiber $\velement \in \vfiber_{\dbasepointc}$. Since $\vchannel$ can be constructed as a pointwise function property, we can construct $\vchannelc$ functions 
\begin{equation}
  \label{eq:construction:nu:fabrication}
  \begin{tikzcd}[col sep=Huge]
    \dfiberc_{\dbasepointc} 
    \arrow[rr, "\vchannelc", color=artist] 
    \arrow[rrrr, "\vchannelc^{\prime\prime}", dashed, bend right, color=artist] &  & 
    \vfiberc_{\dbasepointc}\coloneqq{\dfiberc_{\dbasepointc}^{\prime}} 
    \arrow[rr, "\vchannelc^{\prime}", color=artist] &  & 
    \vfiberc^{\prime}_{\dbasepointc}
  \end{tikzcd}
\end{equation}
such that any new $\vchannel$ can be constructed as a composition of two or more $\vchannel$s.

\subsubsection{Transformer Equivariance}

A pair of transformations $(\vchannel, \vchannel^{\prime})$ are equivariant if 
\begin{equation}
  \dfunct_{\vtotal} \vchannel(\dsection(\dbasepoint)) = \vchannel(\dfunct_{\dtotal}(\dsection(\dbasepoint)))
\end{equation}
for all $\dbasepoint \in \dbase$.  Since the visual sections and data sections are over the same space $\dfunc$

\begin{equation}
  \label{def:construction:nu:validate}
  \begin{tikzcd}[column sep=4em]
    {\dfiberc_{\dbasepointc}^{a}} \times {\dfiberc_{\dbasepointc}^{b}} 
    \arrow[d, "\pi_a"', color=fiber] 
    \arrow[r, "\vchannelc_{ab}", color=artist] 
    \arrow[rr, "\equivc_{ab}", bend left, color=monoid]  & 
    {\vfiberc_{\dbasepointc}^{a}} \times {\vfiberc_{\dbasepointc}^{b}} 
    \arrow[d, "\pi_a", color=fiber] & 
    \measurec_{\dbasepointc}^{ab} 
    \arrow[d, "\measurec\restriction_a", color=set] \\
    \dfiberc_{\dbasepointc}^a 
    \arrow[r, "\vchannelc_{a}", dashed, color=artist] & 
    \vfiberc_{\dbasepointc}^a 
    \arrow[r, "\simeq", dotted]  & 
    \measurec_{\dbasepointc}^a   \\
    {\dfiberc_{\dbasepointc}^{a}} \times {\dfiberc_{\dbasepointc}^{c}} 
    \arrow[u, "\pi_a", color=fiber] 
    \arrow[r, "\vchannelc_{ac}", color=artist] 
    \arrow[rr, "\equivc_{ac}", bend right, color=monoid] & 
    {\vfiberc_{\dbasepointc}^{a}} \times {\vfiberc_{\dbasepointc}^{c}} 
    \arrow[u, "\pi_a"', color=fiber] & 
    \measurec_{\dbasepointc}^{ac} 
    \arrow[u, "\measurec\restriction_a"', color=set]            
    \end{tikzcd}
\end{equation}

\subsection{Compositing Measurable Components into Output}
The compositing function $\vmarkc$ transforms the sections of visual elements $\vsectionc$ into sections of graphics $\gsectionc$.
\begin{equation}
  \vmarkc: \cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}} \rightarrow \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}}
\end{equation}

@@

\subsubsection{Compositor Equivariance}

As mentioned in \autoref{sec:construction:nu}, the distinction between data sections $\dsection$ and visual sections $\vsection$ is semantic rather than structural. Therefore, the equivariance condition on the sheaf map with a change of base space $\vmarkc$ is identical to the one for artists shown in \autoref{eq:artist:equivariance} such that an equivariant \vmark\ must satisify the condition 
\begin{equation}
  \vmarkc(\dfunctc_{\vtotalc}(\vsectionc(\dfunchc_{\vtotalc}(\vindexc(\gbasepointc^{\prime}))))) = \dfunctc_{\gtotalc}(\vmarkc(\vsectionc(\vindexc(\dfunchc_{\gtotalc}(\gbasepointc^{\prime})))))  
\end{equation}
which means that any change to a measurable component input must have a measurably equivalent change in the output.

>>>>

The fiber category is equipped with a bifunctor $\otimes: \mathcal{\dfiber} \times \mathcal{\dfiber} \rightarrow \mathcal{\dfiber}$. The bifunctor provides a method for combining fibers, thereby allowing for fields that contain multityped values. For example, RGB color can either be represented as three fiber fields $\dfiber_{red}, \dfiber_{green}, \dfiber_{blue}$ or a composite fiber field $\dfiber_{red} \otimes \dfiber_{green} \otimes \dfiber_{blue} = \dfiber_{RGB}$.

@@

A sheaf is a presheaf that satisfies the locality and gluing axioms \cite{bakerMathsSheaf}. 
\begin{definition}
\label{def:atct:sheaf:locality}
The locality axiom is that given a union of opensets $\mathscr{\opensetc} = \bigcup\limits_{i\in I} \opensetc_i$ and $\dsectionc^{a}, \dsectionc^{b} \in \sheafc(\mathscr{\opensetc})$,  if $\dsectionc^{a}\restriction_{\opensetc_i} = \dsectionc^{b}\restriction_{\opensetc_i}$ for each $\opensetc_i \in \opensetc$ then $\dsectionc^{a} = \dsectionc^{b}$
\end{definition}
This means that two sections in a sheaf are equal when they evaluate to the same values over all open sets in a collection. 
\begin{definition}
\label{def:atct:sheaf:gluing}
The gluing axiom is that given $\dsectionc^{i} \in \sheafc(\opensetc_i)$ such that ${\dsectionc^{i}}\restriction_{\opensetc_i\cap \opensetc_j} = {\dsectionc^{j}}\restriction_{\opensetc_i \cap \opensetc_j}$ for $\opensetc_i, \opensetc_j \in \mathscr{\opensetc}$, there exists $\dsectionc \in \sheafc(\mathscr{\opensetc})$ such that $\dsectionc\restriction_{\opensetc_i} = {\dsectionc^{i}}$. 
\end{definition}
This means we can construct a section $\dsection$ on the collection of open sets that is the union of sections defined on specific open sets. This expresses an equivalency between a concatenated representation of data and the distributed representation of the same data. In this model, we represent data and graphics as a sheaf; doing so allows us to express that data sources are expected to implement the bookkeeping that a sheaf requires, meaning inclusions, restrictions, locality, and gluing must be preserved.

\subsubsection{Structure}
\label{sec:atct:sheaves:measurments}
We define structure on data as the allowable \textcolor{action}{transformations on the bundle}, which we formalize as actions $\dfuncc$ on sections of that sheaf. Our construction of $\dfunc$ is a generalization of the classification of measurement scales by their mathematical structure \cite{stevensTheoryScalesMeasurement1946,leaFormalizationMeasurementScale, thomasMathematizationNotMeasurement2014}, which often serves as the basis of the equivariance described in \autoref{sec:related-work:equivariance}.

@@

\begin{equation}
  \label{eq:artist:natural_transform:inclusions}
  \begin{tikzcd}
    \opensetc_1 & \cgamma{\opensetc_{1}}{\dtotalc\restruction_{\opensetc_1}} \arrow[dd, "\iota^*"', color=set] \arrow[rr, "\vartistc_{\opensetc_1}", color=artist] &  & \cgamma{\opensetc_{1}}{\vindexpushc\gtotalc\restruction_{\opensetc_1}} \arrow[dd, "\iota^*", color=set] \\
      &  &  &  \\
    \opensetc_{2} \arrow[uu, "\iota", hook, color=base] & \cgamma{\opensetc_{2}}{\dtotalc\restruction_{\opensetc_2}} \arrow[rr, "\vartistc_{\opensetc_2}", color=artist]                        &  & \cgamma{\opensetc_{2}}{\vindexpushc\gtotalc\restruction_{\opensetc_2}}                      
    \end{tikzcd}
\end{equation}
 The diagram in \autoref{eq:artist:natural_transform} shows that restricting a set of outputs of an artist to a set of graphic sections over a subspace is equivalent to restricting the inputs to data sections over the same subspace. By construction, the artist preserves the bookkeeping provided through the use of sheaves and transport functors. Specifically, the artist always takes as input a section over a data subspace and returns a section over the corresponding graphic subspace. Since these are sections of sheaves, they must be implemented such that they satisfy the sheaf axioms (\autoref{def:atct:sheaf:gluing}, \autoref{def:atct:sheaf:locality}) and therefore can be glued together and seperated as needed.

@@

\begin{equation}
  \label{eq:artist:inout:diagram}
  \begin{tikzcd}[row sep=huge]
    \cgamma{\dbasec}{\dtotalc} 
    \arrow[rr, "\vartistc", color=artist] 
    \arrow[d, "\equivc"', color=monoid] &  & 
    \imartist{\gbasec}{\gtotalc} 
    \arrow[d, "render"] 
    \arrow[lld, "\extractmc"', color=monoid, dashed] \\
    {\textcolor{set}{Hom}(\dbasec, \measurec)}  &  & visualization 
    \arrow[ll, "measure",]
    \end{tikzcd}
\end{equation}
so instead we introduce a measurement extraction function $\extractm = measure\circ render$ that decomposes the rendered output into measurable components and returns a map from the data index into the measurement space
\begin{equation}
  \label{eq:artist:actual}
  \extractmc: (\gsectionc \circ \vindexpre) \mapsto (\dbasec \xrightarrow{\extractmc_{\gsectionc}} \measurec)
\end{equation}
such that $\extractmc_{\gsectionc}(\dbasepoint)$ returns a quantification of the measurable components $\measurec_{\dbasepointc}$ of the visual element generated for the data at that point $\gsectionc|_{\vindexprec(\dbasepointc)}$. As shown in \autoref{eq:artist:actual}, the function $\extractmc$ takes in $\gsectionc$ and must evaluate the graphic section on all the graphic space associated with a point $\vindexpre\restriction_{\dbasepoint}$. This is because the measurable aspects, such as thickness or marker shape, refer to the entire visual element for that data point and as such can only be computed on the whole element.

We also introduce a function \equivc\ that maps data to the measurement space directly 
\begin{equation}
\equivc: \dsectionc \mapsto (\dbase \xrightarrow{\equivc_{\dsectionc}} \measurec)
\end{equation}
such that $\equivc_{\dsection}(\dbasepointc)$ is the expected set of measurements $\measure_{\dbasepoint}$. The pair of \textcolor{monoid}{verification functions} (\equivc, \extractmc) can be used to test that the expected encoding $\equivb_{\dsection}$ of the data matches the actual encoding $\extractm_{\gsection}$ 
\begin{equation}
  \label{eq:artist:verification}
    \equivc(\dsectionc)(\dbasepointc) = \extractmc(\vartistc(\dsectionc))(\dbasepointc) = \extractmc(\gsectionc\circ\vindexprec)(\dbasepointc)=\measurec_{\dbasepointc}
\end{equation}

An artist is equivariant when changes to expected and actual are equivariant. As introduced in \autoref{eq:atct:morphism:base}, the base space transformation \dfunch\ is invariant because $\dsection\restriction_{\openset} = \dsection\restriction_{\dfunch(\openset^{\prime})}$. This means that, for all points in the data $\dbasepoint \in \dbase$, the measurement should not change if only the base space is transformed 
\begin{equation}
  \label{eq:atct:equivarance:verify:base}
  \equivc(\dsectionc)(\dfunch(\dbasepointc^{\prime})) = \extractmc(\vartistc(\dsectionc))(\dbasepointc)
\end{equation}
On the other hand, a change in sections \autoref{eq:atct:morphism:fiber} induces an equivalent change in measurements
\begin{equation}
  \label{eq:atct:equivarance:verify:fiber}
  \equivc(\dfunctc(\dsectionc))(\dbasepoint) = \dfunctc_{\measure}(\extractmc(\vartistc(\dsectionc))(\dbasepoint))
\end{equation}
The change in measurements $\dfunctc_{\measure}$ is defined by the developer as the symmetry between data and graphic that the artist is expected to preserve.

@@

\begin{equation}
  \label{eq:artist:plus:verify}
  \begin{split}
  &\vartistc_{a+b}(\dsectionc^{a+b}(\dbasepointc_{a})) \\ 
  & = \vartistc_{a}(\dsectionc^{a}(\dbasepointc_{a})) = \vartistc_{b}(\dsectionc^{b}(\dbasepointc_{b}))
  \end{split}
\end{equation}
 for all $\dbasepointc_{a}, \dbasepointc_{b} \in \dbasec_{a}\bigsqcup\limits_{\dbasec_{c}} \dbasec_{b}$

@@

\begin{equation}
  \begin{split}
    &\extractmc(\vartistc_{a\times b}(\dsectionc^{a\times b}(\dbasepointc)))\restriction_{\dfiberc^{c}}\\ 
    &= 
    \extractmc(\vartistc_{a}(\dsectionc^{a}(\dbasepointc_{a})))\restriction_{\dfiberc^{c}} = \extractmc(\vartistc_{b}(\dsectionc^{b}(\dbasepointc_{b})))\restriction_{\dfiberc^{c}}
  \end{split}
\end{equation}

@@

The circle-line glyph in \autoref{fig:artist:operator} can be generated via a line plot and highlighted points in that plot, which is the addition case of two datasets with the same fields and different continuities. The same circle-line glyph can also be generated via the multiplication case as different fields of the same bundle. As mentioned in \autoref{sec:artist:operator:multiplication}, multiplication is implemented as addition of two copies of a base space. Although we have been discussing the trivial cases of adding observations or adding fields, this merging of artists in datasets can be generalized:
\begin{equation}
  \vartistc(\cgamma{\bigsqcup_{i} \dbasec^{i}}{\bigoplus_{i}\dtotalc^{i}}) \coloneqq \sum_{i}
  \vartisct_{i}(\cgamma{\dbasec^{i}}{\dtotalc^{i}}) 
\end{equation}

@@

\begin{figure}[h!]
  \label{fig:construction:xi}
  \includegraphics[width=1\columnwidth]{deform_retract.png}
  \caption{The graphic base space $\gbase$ is collapsable to the line $\dbase$ such that every band $(\dbasepoint_i, [0,1])$ on $\gbase$ maps to corresponding point $\dbasepoint_i \in \dbase$ and the band $[0,1]$ determines which pixels are filled in to generate the shape of the line.\note{maybe put in a phantom line plot in s}}
\end{figure}
For example, as shown in \autoref{fig:construction:xi}, a line is 1D but is a 2D glyph on a screen; therefore the graphic space $\gbase$ is constructed by multiplying the base space $\dbase$ with an interval $[0,1]$. Because $\gbase$ is collapsible into $\dbase$, every band $(\dbasepoint_i, [0,1])$ corresponds to a point in the base space $\dbasepoint_i \in \dbase$. The first coordinate $\alpha=\dbasepoint_i$ provides a lookup to retrieve the associated visual variables. The second coordinate, which is a point in the interval $\beta=[0,1]$. Together they are a point $\gbasepoint=(\alpha,\beta) \in gbase$ in the graphic base space. This point $\gbasepoint$ is the input into the graphic section $\gsection(\gbasepoint)$ that is used to determine which pixels are colored, which in turn determines the thickness, texture, and color of the line.

\subsection{Encode Data as Measurable Components}
\label{sec:construction:nu}
As introduced in \autoref{fig:construction:flow-chart}, the encoding function $\vchannel$ is a map from data sections to graphic sections
\begin{equation}
\nu: \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \rightarrow \cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}}
\end{equation}
such that the sections project to the same point on the base space $\pi(\dtotal) = \pi(\vchannel(\dtotal))$. A consequence of this property is that $\vchannelc$ can be constructed as a pointwise transformation such that  
\begin{equation}
  \label{eq:constrution:nu}
  \vchannelc: \dfiberc_{\dbasepointc} \rightarrow \vfiberc_{\dbasepoint}
\end{equation}
which means that means that a point in a single data fiber $\delement \in \dfiber_{\dbasepointc}$ can be mapped into a corresponding point in a visual fiber $\velement \in \vfiber_{\dbasepointc}$. This means that an encoding function $\vchannel$ can convert a single record and may not need the whole dataset.

The difference between $\dtotal$ and $\vtotal$ are semantic rather than structural; they are both maps from the topological space \dbase\ to sets of functions that return records of values. This means any $\vtotal$ can be redefined as $\dtotal$, 
\begin{equation}
  \label{eq:construction:nu:fabrication}
  \begin{tikzcd}[col sep=Huge]
    \dfiberc_{\dbasepointc} 
    \arrow[rr, "\vchannelc", color=artist] 
    \arrow[rrrr, "\vchannelc^{\prime\prime}", dashed, bend right, color=artist] &  & 
    \vfiberc_{\dbasepointc}\coloneqq{\dfiberc_{\dbasepointc}^{\prime}} 
    \arrow[rr, "\vchannelc^{\prime}", color=artist] &  & 
    \vfiberc^{\prime}_{\dbasepointc}
  \end{tikzcd}
\end{equation}
which means that, as shown in \autoref{eq:construction:nu:fabrication}, any collection of $\vchannel$ functions can be composed such that they are equivalent to a $\vchannel$ that directly converts the input to the output. This means that a developer has the flexibility to implement either the direct convertor or the building block version (or both) depending on other library concerns. As with artists, $\vchannel$ are maps of sections such that the composition operators defined in \autoref{sec:artist:composition} can also act on transformers $\vchannel$, meaning that encoders can be added and multiplied.

\subsubsection{Encoder Verification}
\label{sec:construction:nu:verification}
A  motivation for constructing an artist with an encoder stage $\vchannel$ is so that the conversion from data to measurable component can be tested separately from the assembly of components into a glyph. 
\begin{equation}
  \label{eq:construction:nu:validate}
  \begin{tikzcd}[column sep=4em]
    {\dfiberc_{\dbasepointc}^{a}} \times {\dfiberc_{\dbasepointc}^{b}} 
    \arrow[d, "\pi_a"', color=fiber] 
    \arrow[r, "\vchannelc_{ab}", color=artist] 
    \arrow[rr, "\equivc_{ab}", bend left, color=monoid]  & 
    {\vfiberc_{\dbasepointc}^{a}} \times {\vfiberc_{\dbasepointc}^{b}} 
    \arrow[d, "\pi_a", color=fiber] & 
    \measurec_{\dbasepointc}^{ab} 
    \arrow[d, "\measurec\restriction_a", color=set] \\
    \dfiberc_{\dbasepointc}^a 
    \arrow[r, "\vchannelc_{a}", dashed, color=artist] & 
    \vfiberc_{\dbasepointc}^a 
    \arrow[r, "\simeq", dotted]  & 
    \measurec_{\dbasepointc}^a   \\
    {\dfiberc_{\dbasepointc}^{a}} \times {\dfiberc_{\dbasepointc}^{c}} 
    \arrow[u, "\pi_a", color=fiber] 
    \arrow[r, "\vchannelc_{ac}", color=artist] 
    \arrow[rr, "\equivc_{ac}", bend right, color=monoid] & 
    {\vfiberc_{\dbasepointc}^{a}} \times {\vfiberc_{\dbasepointc}^{c}} 
    \arrow[u, "\pi_a"', color=fiber] & 
    \measurec_{\dbasepointc}^{ac} 
    \arrow[u, "\measurec\restriction_a"', color=set]            
    \end{tikzcd}
\end{equation}
As shown in \autoref{eq:construction:nu:validate}, an encoder is considered valid if there is an ismorphism between the actual outputted visual component and the expected measurable component encoding. An encoder is consistent if it encodes the same field in the same way even if coming from different data sources.

An encoding function $\vchannelc$ is equivariant if the change in data, as defined in \autoref{sec:atct:sheaves:measurments}, and change in visual components are equivariant. Since $\dtotal$ and $\vtotal$ are over the same base space and are pointwise, the base space change $\dfunch_{\dtotal}$ applies to both sides of the equation 
\begin{equation}
  \vchannelc(\dsectionc_{\dtotalc}(\dfunchc_{\dbasec}(\dbasepointc^{\prime}))) = \vsectionc(\dfunchc_{\dbasec}(\dbasepointc^{\prime}))
\end{equation}
and therefore there should not be a change in encoding. On the other hand, a change in the data values $\dfunct_{\dtotal}$ must have an equivalent change in visual components
\begin{equation}
  \dfunctc_{\vtotal} \vchannelc(\dsectionc(\dbasepoint)) = \vchannelc(\dfunctc_{\dtotalc}(\dsectionc(\dbasepointc)))
\end{equation}
The change in visual components $\dfunct_{\vtotal}$ is dependent both on $\dfunct_{\dtotal}$ and the choice of visual encoding. As mentioned in \autoref{sec:related-work:equivariance}, this is why Bertin and many others since have advocated choosing an encoding that has a structure that matches the data structure\cite{bertinSemiologyGraphicsDiagrams2011a}. For example choosing a quantative colormap to encode quantative data if the $\dfunctc$ operation is scaling, as in \autoref{fig:artist:equivariance}.

\subsection{Composite Measurable Components Into Visual Elements}
As shown in \autoref{fig:construction:flow-chart}, the compositor function $\vmarkc$ transforms the measurable components into properties of a visual element. The compositing function $\vmarkc$ transforms the sections of visual elements $\vsectionc$ into sections of graphics $\gsectionc$.
\begin{equation}
  \vmarkc: \cgamma{\opensetc}{\vtotalc\restriction_{\opensetc}} \rightarrow \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}}
\end{equation}

@@

\subsubsection{Compositor Verification}
\label{sec:construction:q:verification}
An advantage of factoring out encoding and verification, as discussed in \autoref{sec:construction:nu:verification}, is that the responsibility of the compositor can be scoped to translating measurable components into visual elements. 
\begin{equation}
  \label{eq:construction:q:validate}
  \begin{tikzcd}[row sep=huge]
    \cgamma{\opensetc}{(\vtotalc^{a}\times \vtotalc^{b})\restriction_{\opensetc}}  
    \arrow[rr, "\vmarkc_{ab}", color=artist] \arrow[d, "\pi_a"', color=total] &  &  \imartistsub{ab}{\opensetgc}{\gtotalc\restriction_{\opensetgc}} 
    \arrow[d, "\measurec\restriction_a \circ \extractmc_{ab}", color=monoid]  \\
   \cgamma{\opensetc}{\vtotalc^{a}\restriction_{\opensetc}} 
   \arrow[rr, "\simeq", dotted] &  & {\textcolor{set}{Hom}(\opensetc, \measurec^{a})}                                                                             \\
    \cgamma{\opensetc}{(\vtotalc^{a}\times \vtotalc^{c})\restriction_{\opensetc}}  \arrow[rr, "\vmarkc_{ac}", color=artist] \arrow[u, "\pi_a", color=total]  &  &  \imartistsub{ac}{\opensetgc}{\gtotalc\restriction_{\opensetgc}} \arrow[u, "\measurec\restriction_a \circ \extractmc_{ac}"', color=monoid]
   \end{tikzcd}
\end{equation}
As illustrated in \autoref{eq:construction:q:validate}, a compositor is valid if there is an isomorphism between the actual outputted measured visual component and the expected measurable component that is the input. One way of verifying that a compositor is consistent is by verifying that it passes through one encoding even while changing others. For example, when $\vmark_{ab}=\vmark_{ac}$ then the output should differ in the same measurable components as $\vsectionc_{ab}$ and $\vsection_{ac}$.

A compositor function \vmark\ is equivariant if the renderer output changes in a way equivariant to the data transformation defined in \autoref{sec:atct:sheaves:measurments}. This means that a change in base space $\dfunch_{\dtotalc}$ should have an equivalent change in visual element base space. This means that there should be no change in visual measurement
\begin{equation}
  \label{eq:construction:q:verification:base}
  \vsectionc(\dfunchc_{\dbasec}(\dbasepointc^{\prime})) = \extractmc(\vmarkc(\vsectionc)(\dfunch_{\dbasec}(\vindexpre^{\dbasepointc}))) = \measurec_{\dbasepointc}
\end{equation}
As discussed in \autoref{fig:artist:equivariance}, the change in base space may induce a change in locations of measurements relative to each other in the output; this can be verified via checking that all the measurements have not changed relative to the original positions $\measurec_{\dbasepoint} = \measurec_{\dbasepointc^{\prime}}$ and through separate measurable variables that encode holistic data properties, such as orientation or origin.
